Data Predictive Control with Random Forests:\\Theory and Applications

### 2018-01-15 paper/final/intro.tex
intro edits 1

Machine learning and control theory are two foundational but disjoint communities. Machine learning requires data to produce models, and control systems require models to provide stability and performance guarantees to plant operations. Machine learning is widely used for regression or classification, but thus far data-driven models have not been suitable for closed-loop control of physical plants. The challenge now, with using data-driven approaches, is to close the loop for real-time control and decision making.

\subsection{Motivation}
TO IMPROVE $\rightarrow$ \textcolor[rgb]{0,0,1}{Achin and Francesco}

Consider a multivariable dynamical system subject to external disturbances. The first and foremost requirement for making any decision is to obtain the underlying control-oriented predictive model of the system. With a reasonable forecast of the external disturbances, these models should predict the state of the system in the future and thus Model Predictive Control (MPC) can act preemptively to provide a desired system behavior while optimizing a desired performance. In particular, MPC has been proven to be very powerful for multivariable systems in the presence of input and output constraints, and forecast of the disturbances. The caveat is that MPC requires a reasonably accurate physical representation of the system. This makes MPC unsuitable for control of complex plants such as natural gas processing, oil refineries, boilers, manufacturing plants, and buildings where the user expertise, time, and associated sensor costs required to develop a model are very high \cite{Sturzenegger2016,vzavcekova2014}.

There are two main reasons for model complexity. 
(1) The prime contributor is the change in model properties over time. Even if the model is identified once via an expensive route, as the model changes with time, the system identification must be repeated to update the model. Thus, model adaptability or adaptive control is desirable for such systems. 
(2) A secondary reason is the model heterogeneity which further prohibits the use of model-based control. For example, unlike the automobile or the aircraft industry, each building is designed and used in a different way. Therefore, this modeling process must be repeated for every new building. 
Due to aforementioned reasons, the control strategies in such systems are often limited to fuzzy logic rules that are based on best practices.

\paragraph{Example} \textcolor[rgb]{0.00,0.00,1.00}{As a first} example of such modeling complexity, consider the grey-box approach in \cite{Braun2002}. The scope of such approach is to predict the heat transfer rate to the air within the building. In particular, the authors addressed a single zone modeling problem using an equivalent $RC$ model. To this end, five different types of structures are considered in the example: external walls, ceiling/roof, floor, internal walls and windows. Each of these elements are represented using $3$ resistances, $2$ capacitances and $2$ temperatures, except for the windows that are represented only with resistances since they have a negligible energy storage. A model consisting of $8$ states and $9$ inputs, including disturbances such as solar radiation and others, with the instantaneous heat gain to the building air from all surfaces as output, is created for the zone. To have a complete LTI model, $9$ parameters are needed to be estimated via a non-trivial training process, using a collection of information associated with a physical description of the building (see \cite{Braun2002} for more details). Extending this result, for the building we consider in Section \ref{S:realCaseStudy} with $10$ different zones, we would need to consider a model with approximately $80$ states and $90$ parameters to be estimated.

\subsection{Modeling complexity: physics-based vs data-driven}
\textcolor[rgb]{0,0,1}{In this section we provide a more detailed example that quantifies the difference in terms of complexity when modeling a building using physical laws and data.
We compare the 2 approaches to show that the modeling technique used in our solution, i.e. data-driven modeling, eliminates several drawbacks that occur using physics-based models, such as the need to have very good knowledge of the building structure and the materials, time required to build a model, low availability of sensor measurements, etc..
\subsubsection{physics-based modeling example}
This example is taken from \cite{Sturzenegger2016}, where a building located in Allschwil, Switzerland, consisting of 6 floors, is considered. The building total conditioned floor area is of $6000\ m^2$ ca.. The building is modeled as a bilinear system constructed from physical principles. Only the second floor has been modeled, and has been assumed to represent also the others. The modeling process was divided into 4 main points:
\begin{enumerate}
	\item Building geometry and construction data were used together with first-principles laws to derive the following linear model for the building's thermal dynamics:
		\begin{equation}\label{E:RCZoneEq}
			\dot x(t) = A_x x(t) + B_q q(t).
		\end{equation}
		This model describes the behavior of zone, wall, floor and ceiling temperatures.
		Walls, floors and ceilings are considered divided into layers with different features. Therefore each zone was described with an RC network model (see Figure 3-10 in \cite{SturzeneggerTR} for more details), where the capacitances represent the states of the layers and the resistances represent the thermal resistance of the layers.
%		\begin{figure}[t]
%			\begin{center}
%				\includegraphics[width=0.6\textwidth]{figures/RC_net.eps}
%				\caption{RC model of a zone. $q$ represents the external heat fluxes}\label{F:RCnet}
%			\end{center}
%		\end{figure}
		The heat exchange between two adjacent layers, i.e. layer "a" and layer "b", was modeled to be proportional to the temperature difference of the two layers and the corresponding thermal resistance $R$, as follows
		\begin{equation}\label{E:RCHeatExchange}
			\begin{aligned}
				&C_a\dot x_a = \left(x_b(t) - x_a(t)\right)/R\\
				&C_b\dot x_b = \left(x_a(t) - x_b(t)\right)/R,
			\end{aligned}
		\end{equation}
		where $C_a$ and $C_b$ are the heat capacitances of the layers. This is done for each layer of each zone, obtaining the compact model \eqref{E:RCZoneEq}. The thermal parameters were derived from zones geometry and materials data.
	\item External heat fluxes were modeled with a bilinear model of the fluxes which are both direct into the building and indirect through elements and zones:
		\begin{equation}\label{E:RCfluxes}
			q(t) = A_q x(t) + B_{q,u}u(t) + B_{q,d}d(t) + \sum_{i=1}^{n_u}{[\left(B_{q,du,i}d(t) + D_{q,xu,i}x(t)\right)u_i(t)]},
		\end{equation}
		with $u$ the inputs and $d$ the disturbances of the system.
		Equation \eqref{E:RCfluxes} comes from a series of 20 equations analyzed in Section 3.3.1.3 of the Technical Report \cite{SturzeneggerTR}, that we do not report here for the sake of simplicity of the treatise. To model the fluxes several information have been considered in the modeling process:
	\begin{itemize}
		\item heat exchange associated with the building hull (except for windows) was modeled considering a conductive and a radiative part;
		\item  heat flux to each Thermally Activated Building System (TABS, i.e. pipes buried in the concrete slabs of the floors carrying hot/cold water) layer was modeled as a fraction of the total supplied TABS heating and cooling heat fluxes;
		\item heat flux through the windows was considered in three different parts: a radiation that directly acts with the elements in contact with the zone's air; a heat flux due to the conduction through the window; a heat flux due to the absorption of the solar radiation from the window;
		\item internal gains due to occupants, appliances and lighting were modeled as a convecting heat acting in each zone;
		\item effects due to the AHU were considered.
	\end{itemize}
	\item The system was discretized.
	\item The resulting model had approximately 300 states for the only second floor, i.e. temperature of the zones, walls and floors. The output of the system were the zone temperatures.
	Since the performance of the state estimation, that is needed to compute the optimal control inputs using MPC, heavily depends on the number of states, an approximated model with fewer states was needed.
	To this aim, different averaged temperatures of the building facades and the zones was considered, obtaining an approximated model with 35 states. The obtained approximated model was then "suitable for use in MPC" (Section 3.3.1.4 of the Technical Report \cite{SturzeneggerTR}). In particular, after the approximation, the resulting model had the following states, inputs and disturbance variables:
	\begin{itemize}
		\item 5 outputs: averaged room temperature for each group of zones (North, South, West, East, Center). The rooms are equipped with temperature sensors.
		\item 18 inputs: TABS heating heat flux; TABS cooling heat flux; averaged transmitted solar heat flux for each group of zones (North, South, West, East), that was estimated using blinds position measurements; air massflow through the energy recovering mode; air massflow bypassing the energy recovering mode; air massflow through the air cooler; AHU heat coil heat flux; lighting power for the offices for each group of zones (North, South, West, East); radiator heat flux in the corner offices (North, South, West, East).
		\item 7 disturbances: internal gains in the offices and internal gains in non-office zones, which were predicted using standard schedule; ambient temperature and solar radiation on facade (North, South, West, East), whose values were obtained through Kalman filtering, using measurements from the weather station placed on the roof of the building. This filtering is needed to take into account the shadowing of the neighboring buildings.
	\end{itemize}
\end{enumerate}
An EnergyPlus model was built for this system, and model parameters of matrices $A_x$, $B_q$, $A_q$, $B_{q,u}$, $B_{q,d}$, $B_{q,du,i}$, $D_{q,xu,i}$ were derived using geometry and materials data from such EnergyPlus model.
This was a choice of the authors, although in a previous work they used real geometry and materials data from the building. 24 parameters needed to be estimated/taken form a datasheet/computed for the considered zone model.
Although some of the parameters are in common between different zones, the others had to be found independently for each zone.
To derive the EnergyPlus model using the available measurements and to use the same control implemented in the building, the EnergyPlus model was coupled with MATLAB using BCVTB software (see \cite{SturzeneggerTR} for more details). To identify the model a total of 294 MATLAB signals were sent to the EnergyPlus model and a total of 1081 EnergyPlus signals were sent back to Matlab.}

\textcolor[rgb]{0,0,1}{This approach suffers of several drawbacks:
\begin{itemize}
	\item in order to keep the model simple, the heat exchange between layers is modeled with a linear behavior as in \eqref{E:RCHeatExchange}. Hence all the nonlinearities are neglected;
	\item in particular cases, as for example the ventilation energy, a linear model is not good enough to provide a good approximation: for this reason a bilinear model was considered.
	However, also in this case, more complex nonlinearities have been neglected due to the already high complexity of the model.
	For example, in \cite{Sturzenegger2016}, the author says "ideally one would want to formulate the optimization problem in terms of set points and operating modes that can be communicated directly to the BAS.
	However, by limiting the model to a bilinear form, this was not possible";
	\item due to the high modeling complexity, different geometries and different materials of the floors have been neglected supposing they are the same for each floor, although this is almost never true in real cases;  
	\item the resulting model was too complex, so it was further approximated to be "suitable for use in MPC" \cite{SturzeneggerTR};
	\item differently from the building geometry, that could be rebuilt by hand, materials data of the building can be unavailable for different reasons (not provided, lost, changed in time).
	Therefore they need to be estimated, hence further increasing the modeling complexity;
	\item due to the high number of states and variables involved in the modeling framework, a lot of measurements are needed to use the model to predict system's behavior.
	This can be cost expensive due to the amount of sensors needed.
	Furthermore, when some measurements are not directly available from the sensors, observers are needed to provide variables estimations.
	However, observability problems could limit the construction of the observers \cite{Dorf2011MCS}. 
\end{itemize}
}

\subsubsection{data-driven modeling example}
\textcolor[rgb]{0,0,1}{Data-driven modeling helps eliminating all the aforementioned problems...}
TO BE DONE $\rightarrow$ \textcolor[rgb]{0,0,1}{Achin}
\begin{itemize}
	\item show procedure
	\item quantify complexity
\end{itemize}

@@

\paragraph{physics-based vs data-driven modeling example}
TO BE DONE $\rightarrow$ \textcolor[rgb]{0,0,1}{Achin}
\begin{itemize}
	\item compare results
\end{itemize}

\paragraph{Objective} The question now is, can we employ data-driven techniques to reduce the cost of modeling and still exploit the benefits that MPC has to offer? Can we build automatic and data-driven approaches for control purposes that are also adaptive, scalable and interpretable? In this paper we address this problem introducing \textit{Data Predictive Control (DPC)} to bridge the gap between Machine Learning and Predictive Control.
\textcolor[rgb]{0,0,1}{
\subsection{Related work}
This section is divided in 3 parts:
\begin{enumerate}
	\item In Section \ref{SSS:ModelBasedWork} we provide a literature review of the approaches that make use of model-based techniques for energy optimization.
	\item In Section \ref{SSS:DataDrivenWork} we report the main works that use of data-driven methodologies for energy optimization and point out the difference with respect to our contribution.
	\item In Section \ref{SSS:PreviousWork} we underline the differences between the results presented in this work and the ones of our previous papers.
\end{enumerate}
\subsubsection{Model-based related work}\label{SSS:ModelBasedWork}
\subsubsection{Data-driven related work}\label{SSS:DataDrivenWork}}
In the literature, several studies deal with the data-driven control techniques using machine learning algorithms \cite{Hou2013}. For example, Artificial Neural Networks are exploited to create data-driven models to be used as plant simulator in closed-loop with a Supervisory Model Predictive Control in \cite{Afram2017}. In \cite{Macarulla2017} the authors proposed a predictive control strategy based on Neural Networks, for boilers control in buildings, to decide the optimal time to switch-on the plan to guarantee energy saving and thermal comfort. However, the strategy is not easily scalable to different types of plants and does not use optimization theory in the closed-loop scheme. In \cite{Costanzo2016} a reinforcement learning control strategy, called Model-Assisted Batch Reinforcement Learning, is considered to provide data-driven control for the demand response problem in HVAC systems. Reinforcement Learning is a model-free methodology and for this reason it can not be used for on-line optimization in a closed-loop predictive control scheme. In \cite{Ferreira2012} the authors considered a data-driven predictive control based on Neural Networks to guarantee energy saving and  thermal comfort in public buildings. Neural Networks are used in the closed-loop control scheme to determine a thermal comfort index based on parameters that can be measured or estimated, but no  system dynamics with internal state are included into the optimization problem. More papers related to this topic can be found in the literature, but to the best of the authors' knowledge none of them address the problem of including data-driven state models in the optimal predictive control loop, and hence allow to set up a MPC-like optimal control problem, as we do in DPC.

@@

\subsection{Main contribution and paper organization}
TO CLARIFY $\rightarrow$ \textcolor[rgb]{0,0,1}{Achin and Francesco}
\begin{itemize}
	\item Add more details
	\item Emphasize difference between Section 4 and Section 5
	\item Cite appendix
\end{itemize}
TO ADD $\rightarrow$ \textcolor[rgb]{1,0,0}{Tullio}
\begin{itemize}
	\item Add references for weather forecast accuracy
	\item Quantify accuracy for short-term prediction for weather forcast to justify the assumption of perfect knowledge for the disturbance
\end{itemize}
TO ADD $\rightarrow$ \textcolor[rgb]{0,0,1}{Achin and Francesco}
\begin{itemize}
	\item Add subsection in Section 5 with simulation of DPC considering disturbance with added noise and show that we still have good performance
\end{itemize}
%In our previous work \cite{Behl2016,Jain2016,JainACC2017,JainCDC2017}, we introduced the concept of DPC for receding horizon control. In this paper, we extend these results providing the following contributions.
%\begin{enumerate}
%	\item In Section \ref{S:dpc}, we formally present the two following control techniques:
%	\begin{enumerate}
%		\item DPC with regression trees, and
%		\item DPC with random forests.
%	\end{enumerate}
%	\item In Section \ref{S:proof}, we demonstrate the strength of DPC for receding horizon control via one-to-one comparison against a benchmark MPC controller using a bilinear building model whose parameters were identified using experiments on a building in Switzerland. We show that DPC captures 70\% variance in MPC and offers a comparable performance.
%	\item Section \ref{S:casestudy} describes a practical application of DPC for Demand Response, where we apply DPC to a 6 story 22 zone building model in EnergyPlus \cite{Crawley2001} for which model-based control is not economical and practical due to extreme complexity. We show scalability and efficiency of DPC in providing financial incentives to the end-customers bypassing the need for high fidelity models. We observe that DPC provides the desired power reduction with an average error of 3\%.
%	\item In Section \ref{S:realCaseStudy}, we implement DPC on the real data from an off-grid house located in L'Aquila (Italy), to find the optimal ON/OFF scheduling for the heating system in order to save energy while guaranteeing thermal comfort for the occupants. We quantify the total amount of energy saved, with respect to the classical bang-bang controller widely used in houses for temperature control, using an EnergyPlus model built specifically for the house. We show we can perform an energy saving that goes from $25.4\%$, if we guarantee thermal comfort, to $49.2\%$, if we allow very little discomfort in terms of rooms' temperature.
%\end{enumerate}

>>>>

Control-oriented predictive models of an energy system's dynamics and energy consumption, are needed for understanding and improving the overall energy efficiency and operating costs. With a reasonably accurate forecast of future weather and building operating conditions, dynamical models can be used to predict the energy needs of the building over a prediction horizon, as is the case with Model Predictive Control (MPC) \cite{Sturzenegger2016}. However, a major challenge with MPC is in accurately modeling the dynamics of the underlying physical system.
The task is much more complicated and time consuming in case of a large building and often times, it can be even more complex and involved than the controller design itself.
After several years of work on using first principles based models for peak power reduction, and energy optimization for buildings, multiple authors~\cite{Sturzenegger2016, vzavcekova2014} have concluded that the biggest hurdle to mass adoption of intelligent building control is the cost and effort required to capture accurate dynamical models of the buildings.
The user expertise, time, and associated sensor costs required to develop a model of a single building is very high. Thus, the payback period for the upfront hardware and software installation is expected be be too high making MPC an uneconomical choice for energy management.

Why is physics-based modeling hard for complex systems like buildings or what are the reasons of modeling complexities?

\begin{enumerate}
	\item \textbf{Model capture} - a building modeling domain expert typically uses a software tool to create the geometry of a building from the building design and equipment layout plans, add detailed information about material properties, about equipment and operational schedules. There is always a gap between the modeled and the real building and the domain expert must then manually tune the model to match the measured data from the building~\cite{New2012}. 
Moreover, the modeling process also varies from building to building with the construction and types of installed equipment.
Another major downside with physics-based modeling is that enough data is not easily available and guesses for parameter values have to be made, which also requires expert know how.

\item \textbf{Change in model properties} over time - even if the model is identified once via an expensive route as in \cite{Sturzenegger2016}, as the model changes with time, the system identification must be repeated to update the model. Thus, model adaptability or adaptive control is desirable for such systems.

\item \textbf{Model heterogeneity} further prohibits the use of model-based control. For example, unlike the automobile or the aircraft industry, each building is designed and used in a different way. Therefore, this modeling process must be repeated for every new building. 
\end{enumerate}

Due to aforementioned reasons, the control strategies in such systems are often limited to fixed, sometimes ad-hoc, rules that are based on best practices. 
The alternative is to use black-box, or completely data-driven modeling approaches, to obtain a realization of the system's input-output behavior. 
The primary advantage of using data-driven methods is that it has the potential to eliminate the time and effort required to build white and grey box building models. 
Listening to real-time data, from existing systems and interfaces, is far cheaper than unleashing hoards of on-site engineers to physically measure and model the building. Improved building technology and better sensing is fundamentally redefining the opportunities around smart buildings. 
Unprecedented amounts of data from millions of smart meters and thermostats installed in recent years has opened the door for systems engineers and data scientists to analyze and use the insights that data can provide, about the dynamics and power consumption patterns of these systems.

The key question now is: can we employ data-driven techniques to reduce the cost of modeling, and still exploit the benefits that MPC has to offer?
We therefore look for automatic data-driven approaches to control that are also adaptive, scalable and interpretable.
We solve this problem by bridging the gap between Machine Learning and Predictive Control. In particular, in this paper, we present a method based on Random Forests.

%\paragraph{Example} \textcolor[rgb]{0.00,0.00,1.00}{As a first} example of such modeling complexity, consider the grey-box approach in \cite{Braun2002}. The scope of such approach is to predict the heat transfer rate to the air within the building. In particular, the authors addressed a single zone modeling problem using an equivalent $RC$ model. To this end, five different types of structures are considered in the example: external walls, ceiling/roof, floor, internal walls and windows. Each of these elements are represented using $3$ resistances, $2$ capacitances and $2$ temperatures, except for the windows that are represented only with resistances since they have a negligible energy storage. A model consisting of $8$ states and $9$ inputs, including disturbances such as solar radiation and others, with the instantaneous heat gain to the building air from all surfaces as output, is created for the zone. To have a complete LTI model, $9$ parameters are needed to be estimated via a non-trivial training process, using a collection of information associated with a physical description of the building (see \cite{Braun2002} for more details). Extending this result, for the building we consider in Section \ref{S:realCaseStudy} with $10$ different zones, we would need to consider a model with approximately $80$ states and $90$ parameters to be estimated.

\subsection{Example for comparison of modeling complexity: physics-based vs data-driven}
In this section, we provide a detailed example that describes the difference in terms of complexity when we model a building using physical laws and machine learning.
We show that the data-driven modeling eliminates several drawbacks that occur using physics-based models such as the need to have very good knowledge of the building structure and the material properties, time required to build a model, and limited availability of sensor measurements.

The building under consideration is taken from \cite{Sturzenegger2016}. It is located in Allschwil, Switzerland. It has 6 floors of which only the second floor is modeled, and is assumed to represent all other floors. The total air conditioned floor area is $6000\ \mathrm{m^2}$ ca.

\subsubsection{Physics-based modeling}
\label{SSS:physics-based}
We describe the physics-based approach based on an RC network outlined in \cite{Sturzenegger2016}.
The building is modeled as a bilinear system constructed from physical principles, the modeling process consisting of three steps:
\begin{enumerate}
	\item Building geometry and construction data are used together with first-principles to derive the following linear model for the building's thermal dynamics:
		\begin{equation}\label{E:RCZoneEq}
			\dot x(t) = A_x x(t) + B_q q(t).
		\end{equation}
		This model describes the behavior of the zone, wall, floor and ceiling temperatures.
		Walls, floors and ceilings are considered as divided into layers with different features. Therefore, each zone was described with an RC network model (see Figure 3-10 in \cite{SturzeneggerTR}), where the capacitances represent the states of the layers and the resistances represent the thermal resistance of the layers.
%		\begin{figure}[t]
%			\begin{center}
%				\includegraphics[width=0.6\textwidth]{figures/RC_net.eps}
%				\caption{RC model of a zone. $q$ represents the external heat fluxes}\label{F:RCnet}
%			\end{center}
%		\end{figure}
		The heat exchange between two adjacent layers, i.e. layer ``a" and layer ``b", is modeled to be proportional to the temperature difference of the two layers and the corresponding thermal resistance $R$ given by
		\begin{equation}\label{E:RCHeatExchange}
			\begin{aligned}
				&C_a\dot x_a = \frac{x_b(t) - x_a(t)}{R}, \ \text{and}\\
				&C_b\dot x_b =  \frac{x_a(t) - x_b(t)}{R},
			\end{aligned}
		\end{equation}
		where $C_a$ and $C_b$ are the heat capacitances of the layers. This is done for each layer of each zone, obtaining the compact representation given in \eqref{E:RCZoneEq}. The thermal parameters are derived from zones geometry and material properties.
	\item External heat fluxes, modeled as a bilinear model, affect the building directly as well as indirectly through elements and zones:
		\begin{equation}\label{E:RCfluxes}
			q(t) = A_q x(t) + B_{q,u}u(t) + B_{q,d}d(t) + \sum_{i=1}^{n_u}{[\left(B_{q,du,i}d(t) + D_{q,xu,i}x(t)\right)u_i(t)]},
		\end{equation}
		where $u$ are the inputs and $d$ the disturbances of the system.
%		Equation \eqref{E:RCfluxes} comes from a series of 20 equations analyzed in Section 3.3.1.3 of the Technical Report \cite{SturzeneggerTR}, that we do not report here for the sake of simplicity. 
Equation \label{E:RCfluxes} for the heat flux is obtained by modeling
	\begin{enumerate}
		\item heat exchange associated with the building hull (except for windows), both conductive and radiative part,
		\item  heat flux to each thermally activated building system (TABS), i.e. pipes buried in the concrete slabs of the floors carrying hot/cold water,
		\item heat flux through the windows in three different parts: radiation due to elements directly in contact with the zone air, conduction through the window, and absorption of the solar radiation through the window,
		\item convection due to internal gains due to occupants, appliances and lighting, and
		\item the effects of the AHU.
	\end{enumerate}
	\item The resulting system \eqref{E:RCZoneEq} is discretized to obtained a model suitable for finite horizon MPC. This model has approximately 300 states which include temperature of the zones, walls and floors on the second floor. The output of the system are the zone temperatures.
	Since the performance of the state estimation, that is needed to compute the optimal control inputs using MPC, depends on the number of states, an approximated model with fewer states is needed.
	To this aim, different averaged temperatures of the building facades and the zones are considered, obtaining an approximated model with 35 states. This  approximate model is then considered ``suitable for MPC" (see Section 3.3.1.4 in \cite{SturzeneggerTR}). 
\end{enumerate}

	In summary, the approximate model has the following states, inputs and disturbance variables:
	\begin{itemize}
		\item 5 outputs: averaged room temperature for each group of zones (North, South, West, East, Center). The rooms are equipped with temperature sensors.
		\item 18 inputs: TABS heating heat flux, TABS cooling heat flux, averaged transmitted solar heat flux for each group of zones (North, South, West, East) that is estimated using blinds position measurements, air massflow through the energy recovering mode, air massflow bypassing the energy recovering mode, air massflow through the air cooler, AHU heat coil heat flux, lighting power for the offices for each group of zones (North, South, West, East), and radiator heat flux in the corner offices (North, South, West, East).
		\item 7 disturbances: internal gains in the offices and internal gains in non-office zones which are predicted using a standard schedule, ambient temperature and solar radiation on facade (North, South, West, East) whose values were obtained through Kalman filtering using measurements from the weather station placed on the roof of the building. This filtering is needed to take into account the shadowing of the neighboring buildings.
	\end{itemize}
An EnergyPlus model is built for this system, and model parameters of matrices $A_x$, $B_q$, $A_q$, $B_{q,u}$, $B_{q,d}$, $B_{q,du,i}$, $D_{q,xu,i}$ are derived using geometry and materials data from the EnergyPlus model.
This was a choice of the authors; real geometry and materials data can also be used to estimate the model parameters. 
For this particular building, 24 parameters are estimated/taken form a datasheet/computed for the considered zone model.
Although some of the parameters are in common between different zones, the others are found independently for each zone.
To derive the EnergyPlus model using the available measurements and to use the same control implemented in the building, the EnergyPlus model is coupled with MATLAB using BCVTB \cite{Wetter2015}. 
%To identify the model a total of 294 MATLAB signals were sent to the EnergyPlus model and a total of 1081 EnergyPlus signals were sent back to Matlab.
\\

The physics-based approach, although detailed and accurate (in some cases), is cost and time prohibitive. The long and complex modeling procedure that is unique to every building poses a challenge in applying MPC to large scale buildings.

\subsubsection{Data-driven modeling}
The data-driven approach, on the other hand, ignores the physical details of the building operation. Instead, we use machine learning to learn black-box models where our objective is to find the hyperparameters (for a given structure like Random Forests) of a model that best explains the input-output relationship.
\\

The goal with the data-driven modeling is to learn a function map
\begin{equation}
y(t) = f(u(t),\dots,u(t-\delta_u), d(t),\dots, d(t-\delta_d)),
\end{equation}
where \(y\) represents the outputs, \(u\) the inputs and \(d\) the disturbances as defined earlier, and \(\delta\) is a lag variable. If \(y\) is the average room temperature for one of the zones, a nonlinear function \(f\) is more appropriate, for example \(f\) may belong to the class of Neural Networks or Random Forests.
\\

The data-driven modeling in the case of buildings has several advantages over the physical counterpart.
\begin{enumerate}

	\item \textbf{Domain expertise:} This procedure eliminates the need of a building expert, simplifies the modeling complexities described in Section \ref{SSS:physics-based} and ultimately reduces the cost of modeling a building. Further, for a new building, given the historical data from the building, the same process can be repeated to identify a control-oriented model for MPC.

	\item \textbf{Linearity assumption:} During physical modeling in \eqref{E:RCHeatExchange}, in order to keep the model simple, the heat exchange between layers is modeled with a linear behavior. 
	Hence all the nonlinearities were neglected.
	In the case of heat flux in \eqref{E:RCfluxes}, a linear model is not good enough to provide a good approximation. For this reason a bilinear model was considered.
	However, more complex nonlinearities have been neglected due to the already high complexity of the model \cite{Sturzenegger2016}. 
	In data-driven modeling, we can relax these linearity assumptions as nonlinear functions can be learned rather efficiently and accurately.

	\item \textbf{Other assumptions:} Due to the high modeling complexity, physical modeling neglects different geometries and different materials of the floors assuming they are the same for each floor which is never true in reality.
	Data-driven modeling depends on the sensor data directly. 
	Hence, such assumptions never arise.

%	\item differently from the building geometry, that could be rebuilt by hand, materials data of the building can be unavailable for different reasons (not provided, lost, changed in time).

%	Therefore they need to be estimated, hence further increasing the modeling complexity;

	\item \textbf{Sensor cost:} Due to large number of states and variables in the physical modeling, a lot of measurements are needed to use the model to predict the system's behavior.
	This can be cost expensive due to the quantity of sensors needed.
	Moreover, when some measurements are not directly available from the sensors, observers are needed for state estimations.
	However, observability problems can limit the construction of the observers \cite{Dorf2011MCS}. 
	On the other hand, data-driven modeling relies only on direct measurements from the sensors such as thermostats and multimeters reducing the need for new installations.
\end{enumerate}

@@

\textcolor[rgb]{0,0,1}{
\subsection{Related work}
This section is divided in 3 parts:
\begin{enumerate}
	\item In Section \ref{SSS:ModelBasedWork} we provide a literature review of the approaches that make use of model-based techniques for energy optimization.
	\item In Section \ref{SSS:DataDrivenWork} we report the main works that use of data-driven methodologies for energy optimization and point out the difference with respect to our contribution.
	\item In Section \ref{SSS:PreviousWork} we underline the differences between the results presented in this work and the ones of our previous papers.
\end{enumerate}
\subsubsection{Model-based related work}\label{SSS:ModelBasedWork}
\subsubsection{Data-driven related work}\label{SSS:DataDrivenWork}}
In the literature, several studies deal with the data-driven control techniques using machine learning algorithms \cite{Hou2013}. For example, Artificial Neural Networks are exploited to create data-driven models to be used as plant simulator in closed-loop with a Supervisory Model Predictive Control in \cite{Afram2017}. In \cite{Macarulla2017} the authors proposed a predictive control strategy based on Neural Networks, for boilers control in buildings, to decide the optimal time to switch-on the plan to guarantee energy saving and thermal comfort. However, the strategy is not easily scalable to different types of plants and does not use optimization theory in the closed-loop scheme. In \cite{Costanzo2016} a reinforcement learning control strategy, called Model-Assisted Batch Reinforcement Learning, is considered to provide data-driven control for the demand response problem in HVAC systems. Reinforcement Learning is a model-free methodology and for this reason it can not be used for on-line optimization in a closed-loop predictive control scheme. In \cite{Ferreira2012} the authors considered a data-driven predictive control based on Neural Networks to guarantee energy saving and  thermal comfort in public buildings. Neural Networks are used in the closed-loop control scheme to determine a thermal comfort index based on parameters that can be measured or estimated, but no  system dynamics with internal state are included into the optimization problem. More papers related to this topic can be found in the literature, but to the best of the authors' knowledge none of them address the problem of including data-driven state models in the optimal predictive control loop, and hence allow to set up a MPC-like optimal control problem, as we do in DPC.

@@

\subsection{Main contribution and paper organization}
%TO CLARIFY $\rightarrow$ \textcolor[rgb]{0,0,1}{Achin and Francesco}
%\begin{itemize}
%	\item Add more details
%	\item Emphasize difference between Section 4 and Section 5
%	\item Cite appendix
%\end{itemize}
%TO ADD $\rightarrow$ \textcolor[rgb]{1,0,0}{Tullio}
%\begin{itemize}
%	\item Add references for weather forecast accuracy
%	\item Quantify accuracy for short-term prediction for weather forcast to justify the assumption of perfect knowledge for the disturbance
%\end{itemize}
%TO ADD $\rightarrow$ \textcolor[rgb]{0,0,1}{Achin and Francesco}
%\begin{itemize}
%	\item Add subsection in Section 5 with simulation of DPC considering disturbance with added noise and show that we still have good performance
%\end{itemize}
In our previous work \cite{Behl2016,Jain2016,JainACC2017,JainCDC2017}, we introduced the concept of DPC for receding horizon control. In this paper, we extend these results providing the following contributions.
\begin{enumerate}
	\item In Section \ref{S:dpc}, we formally present the two following control techniques:
	\begin{enumerate}
		\item DPC with regression trees, and
		\item DPC with random forests.
	\end{enumerate}
	\item In Section \ref{S:proof}, we demonstrate the strength of DPC for receding horizon control via one-to-one comparison against a benchmark MPC controller using a bilinear building model whose parameters were identified using experiments on a building in Switzerland. We show that DPC captures 70\% variance in MPC and offers a comparable performance.
	\item Section \ref{S:casestudy} describes a practical application of DPC for Demand Response, where we apply DPC to a 6 story 22 zone building model in EnergyPlus \cite{Crawley2001} for which model-based control is not economical and practical due to extreme complexity. We show scalability and efficiency of DPC in providing financial incentives to the end-customers bypassing the need for high fidelity models. We observe that DPC provides the desired power reduction with an average error of 3\%.
	\item In Section \ref{S:realCaseStudy}, we implement DPC on the real data from an off-grid house located in L'Aquila (Italy), to find the optimal ON/OFF scheduling for the heating system in order to save energy while guaranteeing thermal comfort for the occupants. We quantify the total amount of energy saved, with respect to the classical bang-bang controller widely used in houses for temperature control, using an EnergyPlus model built specifically for the house. We show we can perform an energy saving that goes from $25.4\%$, if we guarantee thermal comfort, to $49.2\%$, if we allow very little discomfort in terms of rooms' temperature.
\end{enumerate}
